PLAE enables quick and easy searching across one million ocular transcriptomes

### 2021-10-22 manuscript_00.Rmd
text update

# Distribution of Tissues and Cell Types in scEiaD
```{r, fig.width=8, fig.height=5}
# organ
a <- meta_filter %>%
  mutate(Organ = case_when(Organ == 'Eye' ~ 'Eye',
                           TRUE ~ 'Body'),
         Tissue = gsub('\\.|_', ' ', Tissue),
         Tissue = gsub("RPE$","RPE-Choroid", Tissue),
         Tissue = gsub(" and Aorta","", Tissue)) %>% 
  group_by(Organ, Tissue,organism, study_accession) %>%
  summarise(Count = n()) %>%
  ungroup() %>% 
  ggplot(aes(x=Tissue,y=Count, fill = organism, group = organism)) +
  ylab('Number of Cells') +
  geom_bar(stat='identity', position = 'dodge') +
  cowplot::theme_cowplot() +
  facet_wrap(~Organ, scales = 'free') +
  coord_flip()

# retina cell types
b <- meta_filter %>% filter(Organ == 'Eye') %>% 
  group_by(CellType_predict, study_accession, organism) %>%
  summarise(Count = n()) %>% filter(Count > 10) %>% ungroup() %>% 
  ggplot(aes(x=CellType_predict,y=Count, fill = organism, group = organism)) +
  ylab('Number of Cells') +
  geom_bar(stat='identity', position = 'dodge') +
  cowplot::theme_cowplot() +
  coord_flip()

plot_grid(a,b)
```
# Distribution of ML Retina Cell Types Between Mouse and Human in Pan-Retina Studies
Drop this figure?
```{r, fig.height=5, fig.width=3}
library(tidyverse)
meta_filter <- fst::read_fst('~/data/scEiaD/2021_08_22_meta_filter.fst') %>%
  as_tibble() %>%
  mutate(CellType_predict = case_when(!is.na(TabulaMurisCellType_predict) ~ 'Tabula Muris',
                                      is.na(CellType_predict) ~ 'Unlabelled',
                                      TRUE ~ CellType_predict))

pan_retina_studies <- c('SRP255195','E-MTAB-7316','SRP151023','SRP222001','SRP222958','SRP194595','SRP223254','SRP050054','SRP269634','SRP269635','SRP158528', 'SRP257758','SRP212788')

meta_filter %>% 
  filter(organism %in% c('Mus musculus','Homo sapiens'), 
         Stage == 'Mature', 
         study_accession %in% pan_retina_studies, 
         grepl('Amacr|Bipol|Cone|Epithel|Endoth|Horizon|Muller|Pericy|Ganglion|Rod', CellType_predict)) %>% 
  group_by(study_accession, CellType_predict) %>% 
  summarise(Count = n()) %>% mutate(Percentage = 100 * (Count / sum(Count)))  %>% 
  left_join(meta_filter %>% select(study_accession, Platform, organism) %>% unique()) %>% 
  ggplot(aes(x=organism, y=Percentage)) + 
  geom_boxplot() + 
  coord_flip() + 
  ggbeeswarm::geom_quasirandom(aes(color=study_accession)) + 
  facet_wrap(~CellType_predict, ncol = 1) + 
  cowplot::theme_cowplot()
```
# Introduction
## Many published ocular atlases

## Data not easily available for all

## Some have accompanying apps

## But apps are generally slow and have minimal exploration tooling

@@

Built an awesome app.

@@

## Use scEiaD resource
We built, published [].

## pan mouse and pan human as non-ocular reference
Tabula Muris

>>>>

# (sup?) Table 1
Studies in scEiaD

# Figure 1 Distribution of Tissues and Cell Types in scEiaD
```{r, fig.width=8, fig.height=5}
# organ
a <- meta_filter %>%
  mutate(Organ = case_when(Organ == 'Eye' ~ 'Eye',
                           TRUE ~ 'Body'),
         Tissue = gsub('\\.|_', ' ', Tissue),
         Tissue = gsub("RPE$","RPE-Choroid", Tissue),
         Tissue = gsub(" and Aorta","", Tissue)) %>% 
  filter(Organ == 'Body') %>% 
  group_by(Organ, Tissue,organism, study_accession) %>%
  summarise(Count = n()) %>%
  ungroup() %>% 
  ggplot(aes(x=Tissue,y=Count, fill = organism, group = organism)) +
  ylab('Number of Cells') +
  geom_bar(stat='identity', position = 'dodge') +
  cowplot::theme_cowplot() +
  coord_flip() +
  scale_y_continuous(breaks = c(2500, 5000, 7500, 10000, 12500, 84000))

b <-  meta_filter %>%
  mutate(Organ = case_when(Organ == 'Eye' ~ 'Eye',
                           TRUE ~ 'Body'),
         Tissue = gsub('\\.|_', ' ', Tissue),
         Tissue = gsub("RPE$","RPE-Choroid", Tissue),
         Tissue = gsub(" and Aorta","", Tissue)) %>% 
  filter(Organ == 'Eye') %>% 
  group_by(Organ, Tissue,organism, study_accession) %>%
  summarise(Count = n()) %>%
  ungroup() %>% 
  ggplot(aes(x=Tissue,y=Count, fill = organism, group = organism)) +
  ylab('Number of Cells') +
  geom_bar(stat='identity', position = 'dodge') +
  cowplot::theme_cowplot() +
  facet_wrap(~Organ, scales = 'free') +
  coord_flip()

# retina cell types
c <- meta_filter %>% filter(Organ == 'Eye') %>% 
  group_by(CellType_predict, study_accession, organism) %>%
  summarise(Count = n()) %>% filter(Count > 10) %>% ungroup() %>% 
  ggplot(aes(x=CellType_predict,y=Count, fill = organism, group = organism)) +
  ylab('Number of Cells') +
  geom_bar(stat='identity', position = 'dodge') +
  cowplot::theme_cowplot() +
  coord_flip()

plot_grid(a +  ggbreak::scale_y_break(c(13000, 82000)),b, c)
```

# sup fig 1 
Effect of decontx for Rho between scEiaD v2 and scEiaD v1

<!-- ```{r, fig.height=5, fig.width=3} -->
<!-- # Distribution of ML Retina Cell Types Between Mouse and Human in Pan-Retina Studies -->
<!-- Drop this figure? -->
<!-- Yes -->

<!-- pan_retina_studies <- c('SRP255195','E-MTAB-7316','SRP151023','SRP222001','SRP222958','SRP194595','SRP223254','SRP050054','SRP269634','SRP269635','SRP158528', 'SRP257758','SRP212788') -->

<!-- meta_filter %>%  -->
<!--   filter(organism %in% c('Mus musculus','Homo sapiens'),  -->
<!--          Stage == 'Mature',  -->
<!--          study_accession %in% pan_retina_studies,  -->
<!--          grepl('Amacr|Bipol|Cone|Epithel|Endoth|Horizon|Muller|Pericy|Ganglion|Rod', CellType_predict)) %>%  -->
<!--   group_by(study_accession, CellType_predict) %>%  -->
<!--   summarise(Count = n()) %>% mutate(Percentage = 100 * (Count / sum(Count)))  %>%  -->
<!--   left_join(meta_filter %>% select(study_accession, Platform, organism) %>% unique()) %>%  -->
<!--   ggplot(aes(x=organism, y=Percentage)) +  -->
<!--   geom_boxplot() +  -->
<!--   coord_flip() +  -->
<!--   ggbeeswarm::geom_quasirandom(aes(color=study_accession)) +  -->
<!--   facet_wrap(~CellType_predict, ncol = 1) +  -->
<!--   cowplot::theme_cowplot() -->
<!-- ``` -->
# Introduction
## Many published ocular atlases

## Data difficult to access

@@

Collated XX studies and processed

@@

## Update scEiaD resource
Our first version of scEiaD contained XX studies, 3 species, XX cells, and XX curated cell types. We updated the scEiaD resource for the PLAE web app in several important ways. First, we added two chicken (*Gallus gallus*) studies. Second, we added a human pan-body reference scRNA dataset. Third, we used an *in silico* method to remove background gene contamination. Fourth, we increased the minimum number of detected unique transcripts per cell required for a cell to be retained.

## pan mouse and pan human as non-ocular reference
Tabula Muris and
